noam_data_exploration_v1

### 2024-07-18 noam_data_exploration_v3.Rmd
Worked through turnover, rank shift, and rate change at quadrat level. Rate change is dependent on transformed vs untransformed data. Currently it's based on Euclidean distances, which I don't love. Would like to dig into base code of codyn to see if I can change to Jaccard or Bray-Curtis dissimilarity

```{r appearances}
# examine appearance and disappearances independently 
# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points

appearance <-
  turnover(
  df = long_data,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location",
  metric = "appearance"
)

appearance<- 
  appearance %>%
    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
appearance$Site_habitat <- paste(appearance$Site, appearance$Habitat)
# Add back in stratum column

# A lot of extra spaces

appearance <- appearance %>%
  mutate(across(everything(), str_trim))

str(appearance) # everything became characters

appearance$appearance <- as.numeric(appearance$appearance)
appearance$Year <- as.factor(appearance$Year)
appearance$Site <- as.factor(appearance$Site)
appearance$Habitat <- as.factor(appearance$Habitat)

@@

(appearance_plot <-
  ggplot(data = appearance_means) +
  geom_point(aes(x = Year, y = Appearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Appearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "b. Appearances\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#

disappearance <-
  turnover(
  df = long_data,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Mean_cover",
  replicate.var = "Stratum",
  metric = "disappearance"
)

disappearance<- 
  disappearance %>%
    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
disappearance$Site_habitat <- paste(disappearance$Site, disappearance$Habitat)
# Add back in stratum column

# A lot of extra spaces

disappearance <- disappearance %>%
  mutate(across(everything(), str_trim))

str(disappearance) # everything became characters

disappearance$disappearance <- as.numeric(disappearance$disappearance)
disappearance$Year <- as.factor(disappearance$Year)
disappearance$Site <- as.factor(disappearance$Site)
disappearance$Habitat <- as.factor(disappearance$Habitat)

@@

(disappearance_plot <-
  ggplot(data = disappearance_means) +
  geom_point(aes(x = Year, y = Disappearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Disappearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

# rank shift

rank_shift<-
  rank_shift(
  df = long_data,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Mean_cover",
  replicate.var = "Stratum"
)

rank_shift<- 
  rank_shift %>%
    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
rank_shift$Site_habitat <- paste(rank_shift$Site, rank_shift$Habitat)
# Add back in stratum column

# A lot of extra spaces

rank_shift <- rank_shift %>%
  mutate(across(everything(), str_trim))

str(rank_shift) # everything became characters

rank_shift$MRS <- as.numeric(rank_shift$MRS)
# Create a column with the final year from the returned time.var_pair
rank_shift$Year <- as.numeric(substr(rank_shift$year_pair, 6, 9))
rank_shift$Site <- as.factor(rank_shift$Site)
rank_shift$Habitat <- as.factor(rank_shift$Habitat)

rank_shift_means <-
  rank_shift %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Rank_shift_mean = mean(MRS), 
            Rank_shift_se = sd(MRS)/sqrt(n()))

(rank_shift_plot <-
  ggplot(data = rank_shift_means) +
  geom_point(aes(x = Year, y = Rank_shift_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Rank_shift_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "a. rank_shift\n", y = "Total rank_shift\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )

# rate change
rate_change_table<-
  rate_change(
  df = long_data,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Mean_cover",
  replicate.var = "Stratum"
)
# kable(rate_change_table)

rate_change<-
  rate_change_interval(
  df = long_data,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Mean_cover",
  replicate.var = "Stratum"
)

rate_change<- 
  rate_change %>%
    separate(Stratum, into = c("Site", "Habitat", "Transect"), sep = "-")
rate_change$Site_habitat <- paste(rate_change$Site, rate_change$Habitat)

# A lot of extra spaces

rate_change <- rate_change %>%
  mutate(across(everything(), str_trim))

str(rate_change) # everything became characters

rate_change$distance <- as.numeric(rate_change$distance)
rate_change$interval <- as.numeric(rate_change$interval)
rate_change$Site <- as.factor(rate_change$Site)
rate_change$Habitat <- as.factor(rate_change$Habitat)

ggplot(rate_change, aes(interval, distance, colour = Site, shape = Habitat, group = Site_habitat, linetype = Habitat)) +
  facet_wrap(~Site) + 
  geom_point() +
  scale_colour_manual(values = site_colours) +
  scale_linetype_manual(values = habitat_linetypes) +
  stat_smooth(method = "lm", se = T, size = 2) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site)+
#  labs(y = "Distance\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        ) # I don't think these are linear relationships so the smoother isn't terriibly helpful
```

>>>>

```{r appearances, fig.height = 24, fig.width=16}
# examine appearance and disappearances independently 
# refers to the proportion of species that appeared or disappeared relative to total species pool between two time points

appearance <-
  turnover(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location",
  metric = "appearance"
)

appearance <- merge(turnover_meta, appearance, by = c("Year", "Location"))
appearance$Appearance_trans <- sv_trans(appearance$appearance, appearance)
appearance$Year_factor <- as.factor(appearance$Year)

# model
appearance_turnover_mod1 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site +
                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
                               family = beta_family(), data = appearance)
summary(appearance_turnover_mod1) # low temporal autocorrelation and it's negative
plot(appearance_turnover_res <- simulateResiduals(appearance_turnover_mod1)) # similar residuals problem
Anova(appearance_turnover_mod1) # obviously everything sig

appearance_turnover_mod2 <- glmmTMB(Appearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
                               family = beta_family(), data = appearance)
summary(appearance_turnover_mod2) # 
plot(appearance_turnover_res2 <- simulateResiduals(appearance_turnover_mod2)) #
Anova(appearance_turnover_mod2) 
anova(appearance_turnover_mod2, appearance_turnover_mod1) # models are sig different from each other
AICc(appearance_turnover_mod2, appearance_turnover_mod1) # first model is better but not as drastic as total

plot(allEffects(appearance_turnover_mod1))

@@

(appearance_plot <-
  ggplot(data = appearance_means) +
  geom_point(aes(x = Year, y = Appearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Appearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Appearance_mean - Appearance_se , ymax = Appearance_mean + Appearance_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "b. Appearances\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
#
```

```{r disappearances, fig.height = 24, fig.width=16}
disappearance <-
  turnover(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location",
  metric = "disappearance"
)

disappearance <- merge(turnover_meta, disappearance, by = c("Year", "Location"))
disappearance$Disappearance_trans <- sv_trans(disappearance$disappearance, disappearance)
disappearance$Year_factor <- as.factor(disappearance$Year)

# model
disappearance_turnover_mod1 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site +
                                 (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
                               family = beta_family(), data = disappearance)
summary(disappearance_turnover_mod1) # strong autocorrelation here
plot(disappearance_turnover_res <- simulateResiduals(disappearance_turnover_mod1)) # 
Anova(disappearance_turnover_mod1) #

disappearance_turnover_mod2 <- glmmTMB(Disappearance_trans ~ Year_factor*Habitat*Site + (1 | Site/Transect/Location),
                               family = beta_family(), data = disappearance)
summary(disappearance_turnover_mod2) # 
plot(disappearance_turnover_res2 <- simulateResiduals(disappearance_turnover_mod2)) #
Anova(disappearance_turnover_mod2) 
anova(disappearance_turnover_mod2, disappearance_turnover_mod1) # only case where models are not sig different from each other
AICc(disappearance_turnover_mod2, disappearance_turnover_mod1) # models almost equivalent

plot(allEffects(disappearance_turnover_mod1))

@@

(disappearance_plot <-
  ggplot(data = disappearance_means) +
  geom_point(aes(x = Year, y = Disappearance_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Disappearance_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Disappearance_mean - Disappearance_se , ymax = Disappearance_mean + Disappearance_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "c. Disappearance\n", y = "Total turnover\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
```

```{r mean rank shift, fig.height = 24, fig.width=16}
# rank shift

rank_shift<-
  rank_shift(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location"
)

rank_shift <- merge(turnover_meta, rank_shift, by = c("Location"))

rank_shift_means <-
  rank_shift %>%
  dplyr::group_by(Year, Site, Habitat, Site_habitat) %>%
  dplyr::summarise(Rank_shift_mean = mean(MRS), 
            Rank_shift_se = sd(MRS)/sqrt(n()))
# Looks very weird - not sure what to do with this
(rank_shift_plot <-
  ggplot(data = rank_shift_means) +
  geom_point(aes(x = Year, y = Rank_shift_mean, colour = Site, shape = Habitat, group = Site_habitat), size = 8, stroke = 2) +
  geom_line(aes(x = Year, y = Rank_shift_mean, colour = Site, group = Site_habitat), size = 1.2) +
  geom_errorbar(aes(x = Year, ymin = Rank_shift_mean - Rank_shift_se , ymax = Rank_shift_mean + Rank_shift_se, colour = Site,
                    group = Site_habitat), width = 0.5, size = 1.2) +
  scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site_habitat, nrow = 6)+
  labs(title = "a. rank_shift\n", y = "Total rank_shift\n") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )
    )
```

```{r rate of change, fig.height=15, fig.width=15}
# rate change
long_data_sub$Percent_cover_transformed <- wisconsin(long_data_sub$Percent_cover) # wisconsin transform abundance data

# NOTE: the results here vary considerably if you used transformed vs untransformed data
# Looking into way to change distance matrix from Euclidiean to something else
rate_change_table_transformed<-
  rate_change(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover_transformed",
  replicate.var = "Location"
) # summarizes rate of change for each quadrat over the course of the time series

rate_change_table_transformed <- merge(turnover_meta, rate_change_table_transformed, by = "Location")
hist(rate_change_table_transformed$rate_change)

rate_change_mod1 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_transformed)
summary(rate_change_mod1)
plot(rc_res <- simulateResiduals(rate_change_mod1)) # some deviations, but sig likely due to large n
Anova(rate_change_mod1)
#                Chisq Df Pr(>Chisq)    
# Habitat      187.988  3  < 2.2e-16 ***
# Site          81.209  5  4.686e-16 ***
# Habitat:Site 143.638 15  < 2.2e-16 ***
plot(allEffects(rate_change_mod1))
# Outer 17 has fewest differences
# LTER 1 2 3 changing more than 4 5 6 at outer10
# Backreef change highest at LTER 1 2
# Fringing highest at LTER 1 2 6
emmeans(rate_change_mod1, pairwise ~ Site | Habitat)

# Untransformed
rate_change_table_untransformed<-
  rate_change(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location"
) # summarizes rate of change for each quadrat over the course of the time series

rate_change_table_untransformed <- merge(turnover_meta, rate_change_table_untransformed, by = "Location")
hist(rate_change_table_untransformed$rate_change)

rate_change_mod2 <- glmmTMB(rate_change ~ Habitat*Site + (1|Habitat/Site/Transect), data = rate_change_table_untransformed)
summary(rate_change_mod2)
plot(rc_res <- simulateResiduals(rate_change_mod2)) # as above
Anova(rate_change_mod2)
#                Chisq Df Pr(>Chisq)    
# Habitat       39.224  3  1.556e-08 ***
# Site          17.882  5   0.003098 ** 
# Habitat:Site 123.014 15  < 2.2e-16 ***
plot(allEffects(rate_change_mod1))
# eyeballing it
# Outer 17 scarcely changing regardless of site
# Rate of change similar in backreef between sites
# big variance between sites in outer10 and fringing
# Outer 10: As you go around the island, sites become increasing more "stable" over time (i.e. less shift toward dissimilarity)
# Effectively the opposite trend in Fringing - more dissimilar over time as you go around the island
emmeans(rate_change_mod1, pairwise ~ Site | Habitat)

ggplot(data = rate_change_table_transformed) +
  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
  facet_wrap(~Habitat) +
  scale_colour_manual(values = site_colours) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )

ggplot(data = rate_change_table_untransformed) +
  geom_boxplot(aes(x = Site, y = rate_change, colour = Site) ) +
  geom_jitter(aes(x = Site, y = rate_change, colour = Site), width = 0.2) +
  facet_wrap(~Habitat) +
  scale_colour_manual(values = site_colours) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
        axis.line.y = element_line(colour = "black"),
        axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        axis.line.x = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(size = 32),
        legend.key.size = unit(2, "cm"), # increase legend size
        legend.text = element_text(size = rel(2)), # increase legend text
        legend.title = element_text(size = rel(2.25)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal',
        legend.box.background = element_rect(colour = "black") # make legend horizontal
        )

```{r rate change vis method 2, fig.height=15, fig.width=15}
rate_change_untransformed<-
  rate_change_interval(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover",
  replicate.var = "Location"
)

rate_change_untransformed_distances <- merge(rate_change_untransformed, rate_change_meta, by = "Location")

rate_change_transformed<-
  rate_change_interval(
  df = long_data_sub,
  time.var = "Year",
  species.var = "Taxa",
  abundance.var = "Percent_cover_transformed",
  replicate.var = "Location"
)

rate_change_transformed_distances <- merge(rate_change_transformed, rate_change_meta, by = "Location")

 ggplot(rate_change_untransformed_distances,
        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
   facet_wrap(~Habitat) + 
   geom_point() +
   scale_colour_manual(values = site_colours) +
   scale_linetype_manual(values = habitat_linetypes) +
   stat_smooth(method = "lm", se = T, size = 2) +
   scale_shape_manual(values = habitat_shapes) +
   labs(y = "Untransformed distance\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         )

 ggplot(rate_change_transformed_distances,
        aes(x = interval, y = distance, colour = Site, shape = Habitat) )+
   facet_wrap(~Habitat) + 
   geom_point() +
   scale_colour_manual(values = site_colours) +
   scale_linetype_manual(values = habitat_linetypes) +
   stat_smooth(method = "lm", se = T, size = 2) +
   scale_shape_manual(values = habitat_shapes) +
   labs(y = "Transformed distance\n") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") # make legend horizontal
         ) 
```
